A unified pseudo-$C_\ell$ framework

### 2018-08-31 doc/Paper/main.tex
adding to timing

    \subsection{Curved-sky large-scale structure validation}\label{ssec:validation.lss_full}
      \begin{figure*}
        \centering
        \includegraphics[width=0.99\columnwidth]{val_cl_lss_sph}
        \includegraphics[width=0.99\columnwidth]{val_covar_lss_sph}
        \caption{{\sl Left panel:} average over all simulations in our curved-sky LSS validation suite of the six power spectra between $\delta$, $\gamma_E$ and $\gamma_B$ (solid lines). The analytical prediction for the input signal power spectra are shown as darker dashed lines, and lie mostly under the solid lines (for the non-zero cases $\delta\times\delta$, $\delta\times\gamma_E$ and $\gamma_E\times\gamma_E$). The null power spectra ($\delta\times\gamma_B$, $\gamma_E\times\gamma_B$ and $\gamma_B\times\gamma_B$) are compatible with zero. The middle subplot shows the mean residuals with respect to the input power spectra normalized by the 1$\sigma$ error over the mean (i.e. the scatter over simulations divided by $\sqrt{N_{\rm sim}}$). The residuals are scattered around zero and their fluctuations are well within those expected under the assumption of Gaussian statistics given 1000 independent realizations (see Fig. \ref{fig:lss_val_chi2}). For comparison, the lower panel show the mean relative difference for the power spectra of contaminated maps without contaminant deprojection. {\sl Right panel:} correlation matrix ($C_{ij}/\sqrt{C_{ii}C_{jj}}$) of all measured power spectra estimated from 1000 realizations.} \label{fig:lss_val_cls}
      \end{figure*}
      \begin{figure*}
        \centering
        \includegraphics[width=0.80\textwidth]{val_chi2_lss_sph}
        \caption{Distribution of the $\chi^2$ values found for each of the 1000 simulations in our curved-sky LSS validation suite (blue histograms). Results are shown for all possible cross-correlations, labeled in the top right corner of each subplot. The simulated distributions are compared with the expected $\chi^2$ distribution (solid lines) assuming Gaussian statistics for $N_{\rm dof}=681$ degrees of freedom (denoted by the dashed vertical lines). The dot-dashed vertical lines mark the $\chi^2$ value found for the mean residual with respect to the input power spectrum (see middle panel of the left panel in Fig. \ref{fig:lss_val_cls}). The associated PTE values are $\gtrsim0.1$, and thus no statistically significant bias is found by our validation suite.} \label{fig:lss_val_chi2}
      \end{figure*}
      We generate 1000 realizations of contaminated maps of the galaxy overdensity $\delta$ and cosmic shear $\gamma$ using the models described in Section \ref{sssec:validation.suite.lss} with resolution $N_{\rm side}=1024$. In our fiducial run, we included all contaminants, and used a non-binary mask that traces the shot-noise fluctuations due to the presence of stars and depth variations. For each simulation, we use \nmt~to compute all six possible cross-correlations between the $\delta$ and the $E$ and $B$ modes of $\gamma$ (which we will label $\gamma_{E,B}$), including map-level contaminant deprojection and removal of the noise and deprojection bias in the power spectrum. We use band-powers defined by averaging over groups of $\Delta\ell=3$ multipoles. The typical running time for one realization is $\sim 600$s using 32 \texttt{OpenMP} threads. Subsequent runs are typically faster since we cache the coupling matrices, which are the slowest step in our power-spectra estimation. More details about timing and scalability of \nmt can be found in Appendix \ref{sec:scalability}.

@@

        \caption{{\sl Left:} average over all simulations in our
          curved-sky CMB validation suite of the three spin-2 power
          spectra ($EE$, $EB$ and $BB$, solid lines) using mode
          deprojection and E/B mode purification. The analytical prediction for the input signal power spectra are shown as dashed lines, and line mostly under the solid lines for the non-zero cases ($EE$ and $BB$). The null power spectrum $EB$ is compatible with zero. For comparison, the red dot-dashed line shows the result of not correcting for the mode-deprojection bias when estimating the $B$-mode power spectrum, which leads to a strong bias on small $\ell$. The bottom panel of this figure shows the mean residuals with respect to the input power spectra normalized by the $1\sigma$ error on the mean (i.e. the scatter over simulations divided by $\sqrt{N_{\rm sim}}$). The residuals are scattered around zero with random fluctuations of up to $\sim3\sigma$. {\sl Right:} exact bandpower window functions for the fully accounting for the effects of mode-coupling (solid black) compared with the input top-hat bandpower weights (red). Using the correct window functions is in general important in order to avoid parameter biases. Results are shown for the $BB$ power spectrum.} \label{fig:cmb_val_cls}

@@

    Since in the flat-sky limit $\ell$ is a continuous variable, bandpowers ${\sf B}^{ab}_q$ can be defined as averages over a given interval in $\ell$, $[\ell_q^{\rm min},\ell_q^{\rm max}]$:
    \begin{equation}
      {\sf B}^{ab}_q\equiv \int_{\ell_q^{\rm min}}^{\ell_q^{\rm max}} \frac{d\ell}{\ell_q^{\rm max}-\ell_q^{\rm min}}\,{\sf C}^{ab}_\ell.
    \end{equation}
    The binned coupling matrix is therefore given by:
    \begin{equation}
      \mathcal{M}^{s_as_b}_{qq'}\equiv\int^{\ell_q^{\rm max}}_{\ell_q^{\rm min}} \frac{d\ell}{\ell_q^{\rm max}-\ell_q^{\rm min}}\int^{\ell_{q'}^{\rm max}}_{\ell_{q'}^{\rm min}} d\ell'\,{\sf M}^{s_as_b}_{\ell\ell'}.
    \end{equation}

@@

  \section{Curved-sky validation for different pixel resolutions}\label{app:full_sky_nside}
    We checked that \texttt{NaMaster} returns unbiased power-spectra for all possible values of the HEALPix resolution parameter $N_{\rm side}$. In this appendix we show that we recover the correct power spectra using the inputs from our curved-sky LSS validation suite (see Section \ref{sssec:validation.suite.lss}). In particular we tested $N_{\rm side}$ ranging from 8 to 8192. For the latter, the computational resources needed make it very challenging to generate and analyze a large enough number of simulations to have a good statistical power. However, we generate 100 simulations with $N_{\rm side}=4096$ and show the results in Figure~\ref{fig:validation.fullsky.lss.4096}. We can see that the power spectra are recovered without any noticeable biases in the expected range of validity $\ell \lesssim 2 N_{\rm side}$.

    \begin{figure}
      \centering
      \includegraphics[width=0.99\columnwidth]{val_cl_lss_sph_4096}
      \caption{Input (dashed darker lines) and mean measured (solid semi-transparent lines) power spectra for simulations \texttt{HEALPix} $N_{\rm side}=4096$ using the curved-sky LSS validation suite (see Section \ref{sssec:validation.suite.lss}).}
      \label{fig:validation.fullsky.lss.4096}
    \end{figure}

\bsp

\section{Performance and scalability}
\label{sec:scalability}
     One of the known advantages of the pseudo-$C_{\ell}$ estimator is that it is usually much faster that other known estimators~\citep{2017MNRAS.465.1847E} since the complexity scales as $\mathcal{O}(\ell_{max}^{3})$, where $\ell_{max}$ is the maximum multipole at which the power-spectra are calculated. In order to check how \nmt\, execution time changes with $\ell_{max}$, we compute the curved-sky angular auto and cross-power spectra of a spin-0 and a spin-2 field, with no contaminants. We change the resolution of the \texttt{HEALPix} maps by varying the $N_{side}$ parameter in the range [16, 2048], and we calculate the power-spectra up to $\ell_{max}=3N_{side}$. We use one interactive Cori-Haswell node at the National Energy Research Scientific Computing Center (NERSC). This kind of nodes have two 16-core Intel Xeon (E5-2698 v3) processors at 2.3 GHz, and 128 GB DDR4 of RAM. We executed the code using 32 \texttt{OpenMP} threads. In Figure~\ref{fig:scaling_lmax} we can see the results for this test. We show the results for the auto-power spectra for the spin-0 and spin-2 fields (Data 00, and Data 22, respectively), as well as the results for the cross-power spectra (Data 20). In all cases when we compare with the reference line $\propto \ell_{max}^{3}$ we can see that the code scales as expected. We fixed the proportionality constant for the reference line using the timing of the auto-power of the spin-0 field at $\ell_{max}=96$.  We can also see that for very small $N_{side}$ data handling time starts to be comparable to the pure power-spectra computation time, increasing the expected execution time.

@@

\end{document}

>>>>

    \subsection{Curved-sky large-scale structure validation}\label{ssec:validation.lss_full}
      \begin{figure*}
        \centering
        \includegraphics[width=0.99\columnwidth]{val_cl_lss_sph}
        \includegraphics[width=0.99\columnwidth]{val_covar_lss_sph}
        \caption{{\sl Left panel:} average over all simulations in our curved-sky LSS validation suite of the six power spectra between $\delta$, $\gamma_E$ and $\gamma_B$ (solid lines). The analytical prediction for the input signal power spectra are shown as darker dashed lines, and lie mostly under the solid lines (for the non-zero cases $\delta\times\delta$, $\delta\times\gamma_E$ and $\gamma_E\times\gamma_E$). The null power spectra ($\delta\times\gamma_B$, $\gamma_E\times\gamma_B$ and $\gamma_B\times\gamma_B$) are compatible with zero. The middle subplot shows the mean residuals with respect to the input power spectra normalized by the 1$\sigma$ error over the mean (i.e. the scatter over simulations divided by $\sqrt{N_{\rm sim}}$). The residuals are scattered around zero and their fluctuations are well within those expected under the assumption of Gaussian statistics given 1000 independent realizations (see Fig. \ref{fig:lss_val_chi2}). For comparison, the lower panel show the mean relative difference for the power spectra of contaminated maps without contaminant deprojection. {\sl Right panel:} correlation matrix ($C_{ij}/\sqrt{C_{ii}C_{jj}}$) of all measured power spectra estimated from 1000 realizations.} \label{fig:lss_val_cls}
      \end{figure*}
      \begin{figure*}
        \centering
        \includegraphics[width=0.80\textwidth]{val_chi2_lss_sph}
        \caption{Distribution of the $\chi^2$ values found for each of the 1000 simulations in our curved-sky LSS validation suite (blue histograms). Results are shown for all possible cross-correlations, labeled in the top right corner of each subplot. The simulated distributions are compared with the expected $\chi^2$ distribution (solid lines) assuming Gaussian statistics for $N_{\rm dof}=681$ degrees of freedom (denoted by the dashed vertical lines). The dot-dashed vertical lines mark the $\chi^2$ value found for the mean residual with respect to the input power spectrum (see middle panel of the left panel in Fig. \ref{fig:lss_val_cls}). The associated PTE values are $\gtrsim0.1$, and thus no statistically significant bias is found by our validation suite.} \label{fig:lss_val_chi2}
      \end{figure*}
      We generate 1000 realizations of contaminated maps of the galaxy overdensity $\delta$ and cosmic shear $\gamma$ using the models described in Section \ref{sssec:validation.suite.lss} with resolution $N_{\rm side}=1024$. In our fiducial run, we included all contaminants, and used a non-binary mask that traces the shot-noise fluctuations due to the presence of stars and depth variations. For each simulation, we use \nmt~to compute all six possible cross-correlations between the $\delta$ and the $E$ and $B$ modes of $\gamma$ (which we will label $\gamma_{E,B}$), including map-level contaminant deprojection and removal of the noise and deprojection bias in the power spectrum. We use band-powers defined by averaging over groups of $\Delta\ell=3$ multipoles. The typical running time for one realization is $\sim 600$s using 32 \texttt{OpenMP} threads. Subsequent runs are typically faster since we cache the coupling matrices, which are the slowest step in our power-spectra estimation. More details about timing and scalability of \nmt~can be found in Appendix \ref{app:scalability}.

@@

        \caption{{\sl Left:} average over all simulations in our curved-sky CMB validation suite of the three spin-2 power spectra ($EE$, $EB$ and $BB$, solid lines) using mode deprojection and $B$-mode purification. The analytical prediction for the input signal power spectra are shown as dashed lines, and line mostly under the solid lines for the non-zero cases ($EE$ and $BB$). The null power spectrum $EB$ is compatible with zero. For comparison, the red dot-dashed line shows the result of not correcting for the mode-deprojection bias when estimating the $B$-mode power spectrum, which leads to a strong bias on small $\ell$. The bottom panel of this figure shows the mean residuals with respect to the input power spectra normalized by the $1\sigma$ error on the mean (i.e. the scatter over simulations divided by $\sqrt{N_{\rm sim}}$). The residuals are scattered around zero with random fluctuations of up to $\sim3\sigma$. {\sl Right:} exact bandpower window functions for the fully accounting for the effects of mode-coupling (solid black) compared with the input top-hat bandpower weights (red). Using the correct window functions is in general important in order to avoid parameter biases. Results are shown for the $BB$ power spectrum.} \label{fig:cmb_val_cls}

@@

    Since in the flat-sky limit $\ell$ is a continuous variable, bandpowers ${\sf B}^{ab}_q$ can be defined as averages over a given interval in $\ell$, $[\ell_q^{\rm min},\ell_q^{\rm max}]$:
    \begin{equation}
      {\sf B}^{ab}_q\equiv \int_{\ell_q^{\rm min}}^{\ell_q^{\rm max}} \frac{d\ell}{\ell_q^{\rm max}-\ell_q^{\rm min}}\,{\sf C}^{ab}_\ell.
    \end{equation}
    The binned coupling matrix is therefore given by:
    \begin{equation}
      \mathcal{M}^{s_as_b}_{qq'}\equiv\int^{\ell_q^{\rm max}}_{\ell_q^{\rm min}} \frac{d\ell}{\ell_q^{\rm max}-\ell_q^{\rm min}}\int^{\ell_{q'}^{\rm max}}_{\ell_{q'}^{\rm min}} d\ell'\,{\sf M}^{s_as_b}_{\ell\ell'}.
    \end{equation}
    \begin{figure}
      \centering
      \includegraphics[width=0.99\columnwidth]{val_cl_lss_sph_4096}
      \caption{Input (dashed darker lines) and mean measured (solid semi-transparent lines) power spectra for simulations \texttt{HEALPix} $N_{\rm side}=4096$ using the curved-sky LSS validation suite (see Section \ref{sssec:validation.suite.lss}).}
      \label{fig:validation.fullsky.lss.4096}
    \end{figure}

@@

  \section{Curved-sky validation for different pixel resolutions}\label{app:full_sky_nside}
    We checked that \texttt{NaMaster} returns unbiased power-spectra for all possible values of the HEALPix resolution parameter $N_{\rm side}$. In this appendix we show that we recover the correct power spectra using the inputs from our curved-sky LSS validation suite (see Section \ref{sssec:validation.suite.lss}). In particular we tested $N_{\rm side}$ ranging from 8 to 8192. For the latter, the computational resources needed make it very challenging to generate and analyze a large enough number of simulations to have a good statistical power. However, we generate 100 simulations with $N_{\rm side}=4096$ and show the results in Figure~\ref{fig:validation.fullsky.lss.4096}. We can see that the power spectra are recovered without any noticeable bias in the expected range of validity $\ell \lesssim 2 N_{\rm side}$.

 \section{Performance and scalability}\label{app:scalability}
     One of the known advantages of the pseudo-$C_{\ell}$ estimator is that it is usually much faster that other known estimators~\citep{2017MNRAS.465.1847E} since the complexity scales as $\mathcal{O}(\ell_{\rm max}^{3})$, where $\ell_{\rm max}$ is the maximum multipole at which the power-spectra are calculated. In order to check how \nmt\, execution time changes with $\ell_{max}$, we compute the curved-sky angular auto and cross-power spectra of a spin-0 and a spin-2 field, with no contaminants. We change the resolution of the \texttt{HEALPix} maps by varying the $N_{\rm side}$ parameter in the range [16, 2048], and we calculate the power-spectra up to $\ell_{\rm max}=3N_{\rm side}$. We use one interactive Cori-Haswell node at the National Energy Research Scientific Computing Center (NERSC). This kind of nodes have two 16-core Intel Xeon (E5-2698 v3) processors at 2.3 GHz, and 128 GB DDR4 of RAM. We executed the code using 32 \texttt{OpenMP} threads. In the top panel of Figure~\ref{fig:scaling_lmax} we can see the results for this test. We show the results for the auto-power spectra for the spin-0 and spin-2 fields (Data 00, and Data 22, respectively), as well as the results for the cross-power spectra (Data 02). In all cases we find that the code scales as expected (the solid red line shows the expected $\propto\ell_{\rm max}^3$ behaviour). We can also see that, for very small $N_{side}$, data handling time starts to be comparable to the power-spectrum computation time, increasing the execution time with respect to the expectation.

     Not all parts of the power spectrum estimation take the same amount of time. E.g. even though the scaling with $\ell_{\rm max}$ is the same, the computation of the mode-coupling matrix typically takes longer than a performing single SHT. Computation times can be significantly increased if deprojection over a large number of maps is required or if $E/B$ purification is required. The bottom panel of Figure \ref{fig:scaling_lmax} shows the execution times associated to the most relevant types of operations:

@@

\bsp
\end{document}
